---
title: "Using the gdist package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the gdist package}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The problem

When writing functions that take draws from a certain distribution, it is often difficult to add the flexibility to change distributions.
In the past, my functions, have looked something like this:

```{r past-example, eval=F}
some_function <- function(dist, shape, scale, meanlog, sdlog){
    if(dist=="gamma"){
        x <- rgamma(1000, shape=shape, scale=scale)
    }
    if(dist=="lnorm"){
        x <- rlnorm(1000, meanlog=meanlog, sdlog=sdlog)
    }
    if(!(dist %in% c("gamma", "lnorm"))){
        stop("dist must equal either 'gamma' or 'lnorm'.")
    }
    return(x^2+3*x-1)
}
```

This is both time-consuming and inefficient, as incorporating new distributions means adding more arguments and more front matter to the code.
It only gets worse if there are multiple distributions that need to be drawn from.

```{r past-example2, eval=F}
two_dist_sum <- function(dist1, shape1, scale1, meanlog1, sdlog1,
                         dist2, shape2, scale2, meanlog2, sdlog2){
    if(dist1=="gamma"){
        x1 <- rgamma(1000, shape=shape1, scale=scale1)
    }
    if(dist1=="lnorm"){
        x1 <- rlnorm(1000, meanlog=meanlog1, sdlog=sdlog1)
    }
    if(!(dist1 %in% c("gamma", "lnorm"))){
        stop("dist1 must equal either 'gamma' or 'lnorm'.")
    }
    if(dist2=="gamma"){
        x2 <- rgamma(1000, shape=shape2, scale=scale2)
    }
    if(dist2=="lnorm"){
        x2 <- rlnorm(1000, meanlog=meanlog2, sdlog=sdlog2)
    }
    if(!(dist2 %in% c("gamma", "lnorm"))){
        stop("dist2 must equal either 'gamma' or 'lnorm'.")
    }
    return(x1+x2)
}
```

## The `gdist` solution

Using the `gdist` package, the user can specify their own distributions and supply the arguments for those distributions as needed.

```{r solution, eval=T}
library(gdist)
some_function <- function(dist, ...){
    x <- rdist(dist=dist, 1000, ...)
    return(x^2+3*x-1)
}
y1 <- some_function("gamma", shape=5.8, scale=0.9)
summary(y1)
y2 <- some_function("lnorm", meanlog=1.6, sdlog=0.4)
summary(y2)
```

This allows for more flexibility with less code.

When multiple distributions are required, use the `arg_list` to specify the arguments for each distribution.

```{r solution2, eval=T}
two_dist_sum <- function(dist1, dist1_args, dist2, dist2_args){
    x1 <- rdist(dist1, arg_list=dist1_args)
    x2 <- rdist(dist2, arg_list=dist2_args)
    return(x1+x2)
}
y1 <- two_dist_sum(dist1="lnorm",
                   dist1_args=list(n=1000,
                                   meanlog=1.621,
                                   sdlog=0.418),
                   dist2="lnorm",
                   dist2_args=list(n=1000,
                                   meanlog=1.23,
                                   sdlog=0.79))
summary(y1)
y2 <- two_dist_sum(dist1="gamma",
                   dist1_args=list(n=1000,
                                   shape=5.807,
                                   scale=0.948),
                   dist2="lnorm",
                   dist2_args=list(n=1000,
                                   meanlog=1.23,
                                   sdlog=0.79))
summary(y2)
```

The you may want your users to have the same number of draws from each distribution.
`gdist` allows users to specify an `n` value for the `rdist()` function (as well as `x`, `q`, and `p` for the `ddist()`, `pdist()` and `qdist()` functions, respectively).

```{r solution-n, eval=T}
two_dist_sum_n <- function(n_sims, dist1, dist1_args, dist2, dist2_args){
    x1 <- rdist(dist1, n=n_sims, arg_list=dist1_args)
    x2 <- rdist(dist2, n=n_sims, arg_list=dist2_args)
    return(x1+x2)
}
y1 <- two_dist_sum_n(n_sims=1000,
                     dist1="lnorm",
                     dist1_args=list(meanlog=1.621,
                                     sdlog=0.418),
                     dist2="lnorm",
                     dist2_args=list(meanlog=1.23,
                                     sdlog=0.79))
summary(y1)
y2 <- two_dist_sum_n(n_sims=1000,
                     dist1="gamma",
                     dist1_args=list(shape=5.807,
                                     scale=0.948),
                     dist2="lnorm",
                     dist2_args=list(meanlog=1.23,
                                     sdlog=0.79))
summary(y2)
```

If you don't know the abbreviated name for the distribution, you can supply the full name of the distribution and -- if it is in the `stats` package -- the functions will look it up for you when `lookup_dist=T`.
To learn the abbreviated name of the distribution, set `lookup_verbose=T`.

```{r lookup}
rdist(dist="Log-Normal", n=1, meanlog=1.23, sdlog=0.79,
      lookup_dist=T, lookup_verbose=T)
```

Use `data(dist_lookup_table)` to see a list of the distributions from the `stats` package and their abbreviations.
